Tables

### 2018-06-04 code/figures.rmd
Double-checked computations for table, cleaned up figures file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load data and packages}
library(readr)
library(dplyr)
library(kableExtra)
library(purrr)
library(metafor)

@@

I have four different types of effects to transform to raw correlations:
- mean difference + SE
- mean difference + SD
- RD
- OR

Should I first run MA and then transform effect sizes? ->Not sure how how to treat the raw mean differences, since I only get the MA-estimate estimate and the SE.  More straightforward to convert to common metric first. **Kolden + Ihle give formulas.** From book Cooper & Hedges, 1994.

```{r transformation_functions}
#For raw mean difference + SE
transform_SE <- function(ES, SE, n1, n2){ #ES -> SDp based on meta-analysis lecture 02
  sdpooled <- sqrt(SE^2 / (1 / n1 + 1/n2))
  d <- ES / sdpooled
  corr <- (n1 + n2)^2 / (n1*n2) #correction factor if unequal sample sizes
  r <- d / sqrt(d^2 + corr)
  data.frame(r = r, n = n1 + n2)
}

@@

  # data.frame(b = fit$b[[1]], I2 = fit$I2) #estimate out
  transformed
}
```

Transform effect sizes to raw correlations
```{r transform_df}
res <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(transform_MA, .id = "effect")

res %>% filter(!is.na(r) & r > 1) #checks
res %>% filter(is.na(r)) #Checks
```

Seems okay. Metafor takes raw correlations as input for Fisher's z transformed rma.

```{r MA_df}
summarizer <- function(x){
  fit <- rma(measure = "ZCOR", ri = x$r, ni = x$n, test = "knha", data = x) #takes raw correlation as input
  data.frame(z = fit$b[[1]], I2 = fit$I2, H2trunc = fit$H2, H2 = fit$QE / fit$k) #estimate out
}
#Alternative version of calculating I2 does not make conceptual sense because gives negative values. Alternative method for calculating H2 provides information also on excessive homogeneity (i.e, less variability than expected by chance) for those interested. (that is, does not have a lower limit of 1)

res2 <- res %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(summarizer, .id = "effect") %>% 
  mutate(z = abs(z)) #We are interested in effect size rather than direction
```

Above seems to do what it should.

```{r plot_res}
library(ggplot2)

@@

# ggsave("H2_effect_size.png")
library(cowplot)

effect_het <- plot_grid(I2, H2, nrow = 2, labels = c("A", "B"), align = "v")
# save_plot("effect_het.png", effect_het,
          # nrow = 2, base_aspect_ratio = 1.3)

@@

Note that Fisher's z are the absolute z-scores, because we are interested in effect size rather than direction.

```{r labelled_plot}
library(ggplot2)

ggplot(res2, aes(x = z, y = I2, label = effect)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_text(alpha = 0.3)
```

Qs: Should I keep the I2 based on the original scales, but transform the effect sizes and plot that instead (seems not unreasonable since I2 depends on effect type). -> This seems to make little difference, see below.

```{r}
est_heterogen_smd_raw <- function(x){
  if(any(x[, "effect_type"] == "Risk difference")){ #without the 'any' we will get warnings because we apply 'if' to a vector

    fit <- rma(measure = "RD", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2,  n1i = ntreatment, n2i = ncontrol, test = "knha", data = x)

    fit <-  rma(yi = effect_size, sei = outcome_c2, test = "knha",  data = x)

  } else if(any(x[, "effect_type"] == "Raw mean difference")){

    fit <- rma(measure = "MD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol,test = "knha", data = x)

    fit <- rma(measure = "SMD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol,test = "knha", data = x)

    fit <- rma(measure = "COR", ri = effect_size, ni = Ntotal, test = "knha", data = x)

  } else{ #for all two-group count effects

    fit <- rma(measure = "OR2DL", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2, #standardized, ML1 used 'OR2D' but that = 'OR2DL'
               n1i = ntreatment, n2i = ncontrol, test = "knha", data = x)

  I2_1 <- confint(fit)$random[3, ] #Gives us the I2 estimate and its confidence interval
  data.frame(eff_size = fit$b[[1]], #effect size (point estimate)
             p = fit$pval,
             s_I2 = I2_1[1], s_ci.lb = I2_1[2], s_ci.ub = I2_1[3]) #I2 + CI
}

#Apply function to data
res3 <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") %>% 
  select(effect, s_I2) %>% 
  left_join(res2, .)

```{r}
ggplot(res3, aes(x = z, y = s_I2)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

summary(lm(s_I2 ~ z, data = res3)) #b = 68.23, R2 = 0.67
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.asp = 1.3)
```

```{r load data and packages}
library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data transformation
library(metafor) #To run meta-analyses
library(ggplot2) #Plot data
library(cowplot) #Combine several plots into one

@@

```{r transformation_functions}
#Formulas found in: 
#Borenstein, M. (2009). Effect sizes for continuous data. In H. Cooper, L. V. Hedges, & J. C. Valentine (Eds.), The handbook of research synthesis and meta-analysis (2nd ed., pp. 221-235). New York: Russell Sage Foundation.

#For raw mean difference + SE
transform_SE <- function(ES, SE, n1, n2){ 
  sdpooled <- sqrt(SE^2 / (1 / n1 + 1/n2)) #Borenstein, M. (2009). In Cooper & Hedges. p. 224
  d <- ES / sdpooled
  corr <- (n1 + n2)^2 / (n1*n2) #correction factor if unequal sample sizes. p.234
  r <- d / sqrt(d^2 + corr) #p.234
  data.frame(r = r, n = n1 + n2)
}

@@

  # data.frame(b = fit$b[[1]], I2 = fit$I2) #estimate out
  transformed
}

```{r transform_df}
#Transform effect sizes to raw correlations
res <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(transform_MA, .id = "effect") #apply function and transform back from list to dataframe

```{r MA_df}
summarizer <- function(x){
  fit <- rma(measure = "ZCOR", ri = x$r, ni = x$n, data = x) #takes raw correlation as input
  data.frame(z = fit$b[[1]], I2 = fit$I2, H2trunc = fit$H2, H2 = fit$QE / (fit$k - 1)) #estimates out
}
#Default method of metafor for calculating H2 is truncated at one. Alternative method for calculating H2 provides information also on excessive homogeneity, i.e, less variability than expected by chance (that is, does not have a lower limit of 1). This method computes H2 as originally specified by Higgins & Thompson instead of using the more general method of metafor (see ?print.rma.uni).

#Higgins, J., & Thompson, S. G. (2002). Quantifying heterogeneity in a meta-analysis. Statistics in medicine, 21(11), 1539-1558.

res2 <- res %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(summarizer, .id = "effect") %>% 
  mutate(z = abs(z)) #Absolute values since we are interested in effect size rather than direction
```

```{r plot_res}

@@

# ggsave("H2_effect_size.png")

effect_het <- plot_grid(I2, H2, nrow = 2, labels = c("A", "B"), align = "v") #combine plots
effect_het
# save_plot("effect_het.png", effect_het,
          # nrow = 2, base_aspect_ratio = 1.3)

@@

```{r labelled_plot, fig.asp= 0.618}
ggplot(res2, aes(x = z, y = I2, label = effect)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  xlab("Fisher's z") +
  coord_cartesian(ylim = c(0, 100)) +
  theme_classic() +
  geom_text(alpha = 0.5)
```

### 2018-06-04 code/tables.rmd
Double-checked computations for table, cleaned up figures file

    fit <- rma(measure = "RD", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2,  n1i = ntreatment, n2i = ncontrol, test = "knha", data = x)

@@

    fit <-  rma(yi = effect_size, sei = outcome_c2, test = "knha",  data = x)

  } else if(any(x[, "effect_type"] == "Raw mean difference")){

    fit <- rma(measure = "MD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol,test = "knha", data = x)

@@

    fit <- rma(measure = "SMD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol,test = "knha", data = x)

  } else if(any(x[, "effect_type"] == "r")){

    fit <- rma(measure = "COR", ri = effect_size, ni = Ntotal, test = "knha", data = x)

@@

    fit <- rma(measure = "OR2DL", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2, #standardized, ML1 used 'OR2D' but that = 'OR2DL'
               n1i = ntreatment, n2i = ncontrol, test = "knha", data = x)

>>>>

    fit <- rma(measure = "RD", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2,  n1i = ntreatment, n2i = ncontrol,  data = x)

@@

    fit <-  rma(yi = effect_size, sei = outcome_c2,  data = x)

  } else if(any(x[, "effect_type"] == "Raw mean difference")){
    if(any(x[, "rs"] %in% c("RRR5", "RRR7"))){ #For these us the Knapp and Hartung adjustment of standard errors

      fit <- rma(measure = "MD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol, test ="knha", data = x)
    } else{
      fit <- rma(measure = "MD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol, data = x)
    }

@@

    fit <- rma(measure = "SMD", m1i = outcome_t1, m2i = outcome_c1, sd1i = outcome_t2, sd2i = outcome_c2, n1i = ntreatment, n2i = ncontrol, data = x)

  } else if(any(x[, "effect_type"] == "r")){
    if(any(x[, "rs"] == "ML1")){
      fit <- rma(measure = "COR", ri = effect_size, ni = Ntotal,  data = x)
    } else{ #for ML3
      fit <- rma(measure = "UCOR", ri = effect_size, ni = Ntotal,  data = x, vtype = "UB")
      #Note that ML3 specifies measure = "COR" with vtype = "UB", but that this is specified as above in the latest metafor 
    }

@@

    fit <- rma(measure = "OR2DL", ai = outcome_t1, bi = outcome_t2, ci = outcome_c1, di = outcome_c2, 
               n1i = ntreatment, n2i = ncontrol,  data = x)
    #standardized, ML1 used 'OR2D' which = 'OR2DL' in latest version of metafor
